---
title: "Project Machine Learning"
author: "Kassim"
date: "08/02/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(12367)
library(tidyverse)
require(lubridate)
library(caret)

url.training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url.testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read_csv(file = url.training, na = c("NA","#DIV/0!",""))
testing <- read_csv(file = url.testing, na =c("NA","#DIV/0!",""))



```
##Learning about the dataset and adjusting some variavles
###I got some errors when runing the train method because has some missings values. So I kept only class column and the coluns that has numeric values. Obs: Removed all calculated columns, they have to much NA. 
```{r}

  ##removed all calculated columns, they have to much NA. 
  training <- training[, !grepl("total_|kurtosis_|skewness_|max_|min_|amplitude_|var_|avg_|stddev_|timestamp|window|X1|user_name", colnames(training))]

  #transform classe in factor
  training$classe <- as.factor(training$classe)
  print('original file length')
  nrow(training)
  #since the datasource has 3 rows empty, I choose to remove all rows that has NA
  training <- training[complete.cases(training),]
  print('usinging complete.cases in training file')
  nrow(training)
  summary(training)
  
  ##Repeat with test source
  testing <- testing[, !grepl("total_|kurtosis_|skewness_|max_|min_|amplitude_|var_|avg_|stddev_|timestamp|window|X1|user_name", colnames(testing))]
```
##how many classes exist and how much of each ?
```{r}
  table(training$classe)
  prop.table(table(training$classe))
```


#Training the model
##Create a partition to train and to test
##Create 5 folds to train
```{r}
  partition <- createDataPartition(training$classe, p= 0.6, list = F)
  train <- training[partition,]
  test <- training[-partition,] 

  
  traincontrl <- trainControl( method = "repeatedcv", number = 5, repeats = 3)
  
  model.rf <- train(classe ~ ., data = train, trControl = traincontrl, method = "rf")
  model.gbm <- train(classe ~ ., data = train, trControl = traincontrl, method = "gbm")
  
  model.rf 
  model.gbm  
  
  plot(model.rf)
  plot(model.gbm)
```
#Because GBM train resulted in 95.8% of accuracy and random forest had 98.8%, I choose rf model
```{r}

  model.gbm.trained <- predict(model.gbm, newdata=train[,-49])
  confusionMatrix(model.gbm.trained, train$classe)
  
  
  
  model.rf.trained <- predict(model.rf, newdata=train[,-49])
  confusionMatrix(model.rf.trained, train$classe)
```
###And the confusion matrix of RF predict got 100% of accuracy.
##Lets try with test data source

```{r}
  resulttesting <- predict(model.rf, testing, method = "class")

  table(resulttesting)
  prop.table(table(resulttesting))
```



